Learning Auditory Similarities Between Instrumental Playing Techniques

### 2020-05-25 paper/paper.tex
commentaires
remover overset t in \ast
commit ML's conclusion

The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain. \ml{even if research about its notation is of interest}
Instead, it involves a cognitive process which arises from the subjective experience of listening \cite{erickson1975book}.
The task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
Rather than merely discriminating instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters---duration, pitch, and intensity---being equal \cite{thoret2018jasa}.

@@

Let us recall the full definition of joint time--frequency scattering (see \lnameref{sec:methods} section):
\begin{equation}
\mathbf{S_2}\boldsymbol{x}(t,\lambda,\alpha,\beta) =
\left(
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\circledast
\boldsymbol{\Psi}_{\alpha,\beta}
\Big\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_{F})
\right)(t,\lambda),
\label{eq:joint-scattering}
\end{equation}
where the $\boldsymbol{\psi}_{\lambda}$ denotes a Morlet wavelet of center frequency $\lambda$ and resp. $\boldsymbol{\phi}_T$ denotes a Gaussian low-pass filter of width $T$.
Besides the \emph{joint} time--frequency scattering transform, the generative grammar of scattering transforms \cite{lostanlen2019chapter} also encompasses the \emph{separable} time--frequency scattering transform:
\begin{equation}
\mathbf{S_2^{\mathrm{sep}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) =
\left(
\bigg\vert
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\ast
\boldsymbol{\psi}_{\alpha}
\Big\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\psi}_{\beta}\big)
\bigg\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_F\big)
\right)(t, \lambda),
\label{eq:separable-scattering}
\end{equation}
where the wavelet $\boldsymbol{\psi}_{\lambda}$ has a quality factor of $Q=12$ whereas the wavelets $\boldsymbol{\psi}_{\alpha}$ and $\boldsymbol{\psi}_{\beta}$ have a quality factor of one.
Previous publications have successfully applied separable time--frequency scattering in order to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques involving periodic modulations \cite{wang2019ismir}.

@@

That being said, the above model is general enough to encompass a wide variety of recordings, not just instrumental sounds.
Indeed, \ml{we have strong assumptions} that the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed. \ml{Though, future work should demonstrate that they} work for more general classes of sounds.
To extend the model, it is only necessary to retrain the LMNN weighting matrix by supplying it with new cluster assignments.
These can again be obtained by performing a new clustering experiment with one or more human subjects.

>>>>

% Le numerique permet d'approcher la question de la notation de facon implicite
% Cela ne retire pas l'interet de l'etude des graphemes ou du texte (glissando vs slide)
The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain. \ml{even if research about its notation is of interest}
Instead, it involves a cognitive process which arises from the subjective experience of listening \cite{erickson1975book}.
The task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
Rather than merely discriminating instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters---duration, pitch, and intensity---being equal \cite{thoret2018jasa}.

@@

Let us recall the full definition of joint time--frequency scattering (see \lnameref{sec:methods} section):
\begin{equation}
\mathbf{S_2}\boldsymbol{x}(t,\lambda,\alpha,\beta) =
\left(
\Big\vert
\big\vert
\boldsymbol{x} \ast \boldsymbol{\psi}_{\lambda}
\big\vert
\circledast
\boldsymbol{\Psi}_{\alpha,\beta}
\Big\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_{F})
\right)(t,\lambda),
\label{eq:joint-scattering}
\end{equation}
where the $\boldsymbol{\psi}_{\lambda}$ denotes a Morlet wavelet of center frequency $\lambda$ and resp. $\boldsymbol{\phi}_T$ denotes a Gaussian low-pass filter of width $T$.
Besides the \emph{joint} time--frequency scattering transform, the generative grammar of scattering transforms \cite{lostanlen2019chapter} also encompasses the \emph{separable} time--frequency scattering transform:
\begin{equation}
\mathbf{S_2^{\mathrm{sep}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) =
\left(
\bigg\vert
\Big\vert
\big\vert
\boldsymbol{x} \ast \boldsymbol{\psi}_{\lambda}
\big\vert
\ast
\boldsymbol{\psi}_{\alpha}
\Big\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\psi}_{\beta}\big)
\bigg\vert
\circledast
\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_F\big)
\right)(t, \lambda),
\label{eq:separable-scattering}
\end{equation}
where the wavelet $\boldsymbol{\psi}_{\lambda}$ has a quality factor of $Q=12$ whereas the wavelets $\boldsymbol{\psi}_{\alpha}$ and $\boldsymbol{\psi}_{\beta}$ have a quality factor of one.
Previous publications have successfully applied separable time--frequency scattering in order to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques involving periodic modulations \cite{wang2019ismir}.

@@

That being said, the above model is general enough to encompass a wide variety of recordings, not just instrumental sounds.
Indeed, we have strong assumptions that the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed. Though, future work should demonstrate that they work for more general classes of sounds.
To extend the model, it is only necessary to retrain the LMNN weighting matrix by supplying it with new cluster assignments.
These can again be obtained by performing a new clustering experiment with one or more human subjects.
